%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Flowlog NSDI submission
% PAGE LIMIT: 12 pages, NOT including references
% 10 point, 2 column
% single blind (don't anon)
% *** BEWARE SEPARATE ABSTRACT SUBMISSION DEADLINE ***
% ABSTRACT SUBMISSION 9/20
% PAPER SUBMISSION 9/28

\documentclass[letterpaper,twocolumn,10pt]{article}
% ``As of 2010, USENIX does not require endnotes 
%\usepackage{usenix,epsfig,endnotes}
\usepackage{usenix,epsfig}

\usepackage{fancyvrb}

\include{tim-preamble}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf A Balance of Power: etc etc}

\author{
{\rm Your N.\ Here}\\
Your Institution
\and
{\rm Second Name}\\
Second Institution
 \and
 {\rm Name}\\
Name Institution
} % end author


\maketitle

\begin{abstract}
stuff
\end{abstract}

\section{Introduction}

similar story to the hotsdn submission. 

\subsection{Contributions}

context: beyond hotsdn paper. external code, external events, far
richer examples, new verification work, downloadable
implementation (can we manage this? surely will make it public if
accepted at least)

\section{Examples}

\begin{itemize}
  \item{Stolen Laptop}
  \item{ARP Cache}
  \item{NIB}
  \item{Forwarding (smart macl/span tree/shortest paths?)}
  \item{Zero config}

  \item{[Maybe?] bandwidth res plus sch downtime. May not have
time. Discuss AF.}

\end{itemize}

\section{Verification}

PVS? Select some examples, + properties to verify. 

(can we verify liveness/safety/etc. in pvs?)

\section{Flowlog, formally}

division of the 2 languages. core syntax. semantics. syntactic sugar (+/-/etc)

\section{Discussion}

  \subsection{Declarative, Rule-based (others have done, too)}

  \subsection{???}

  \subsection{Preliminary Benchmarks}

  \subsection{Proactivity and Optimization}

\section{Related Work}

need to flesh out a lot of hotsdn's rw also:

- ndlog/overlog: verification AND rule based.

- dedalus (especially if we use it in our semantics)

- op sem decl networking. [other papers by this group?]. may not
  go in relwork but get cited in body: ambiguities we avoid.

- other networking refs. [fill]

\section{Conclusion}



Future work: more analyses, synthesis, change-impact. Scaling as
discussed. 



{\footnotesize \bibliographystyle{acm}
\bibliography{tn}}

\end{document}
